---
title: "Application of bcimodel to East Africa with HER2 extension"
author: "Roman Gulati"
date: "`r format(Sys.time(), "%B %d, %Y")`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Application to East Africa with HER2 extension"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", echo=TRUE)
```

## Overview
This vignette reproduces application of the [bcimodel]() to Uganda and extends this application to project expected impacts of mammography and aggressive treatments for appropriate cases, including effects of trastuzumab for HER2+ cancers.

## Create template for treatment distributions
This is a simple function to format and initialize the treatment distribution model input.
```{r create-treatment-distribution}
library(bcimodel)
create_treatment_distribution <- function(nh,
                                          pol,
                                          treat_hrs=c('None'=1, 'Tamoxifen'=0.7),
                                          treat_probs=c('None'=1, 'Tamoxifen'=0),
                                          scale_by_subgroup=FALSE){
    ssids <- with(nh, paste(stage, subgroup, sep='.'))
    tset <- expand.grid(txHR=treat_hrs,
                        SSid=ssids,
                        stringsAsFactors=FALSE)
    tset <- with(tset, data.frame(SSno=as.integer(factor(SSid, levels=unique(SSid))),
                                  SSid,
                                  txSSno=seq_len(nrow(tset)),
                                  txSSid=names(treat_hrs),
                                  txHR,
                                  stringsAsFactors=FALSE))
    tprobs <- data.frame(txSSid=names(treat_probs),
                         treat_probs,
                         row.names=NULL,
                         stringsAsFactors=FALSE)
    tprobs <- reshape::cast(merge(tprobs, data.frame(policy=pol$id)),
                            txSSid~policy,
                            value='treat_probs')
    merged <- merge(tset, tprobs, by='txSSid')
    merged <- transform(merged, subgroup=sub('^[^.]*[.](.*)$', '\\1', SSid))
    if(scale_by_subgroup){
        sgids <- plyr::ddply(nh, 'subgroup', plyr::summarize, prop=sum(prop))
        merged <- merge(merged, sgids, by='subgroup')
        for(policy in as.character(pol$id))
            merged[[policy]] <- merged[[policy]]*merged[['prop']]
    }
    merged <- reshape::sort_df(merged, vars='txSSno')
    merged <- merged[, c('SSno', 'SSid', 'txSSno', 'txSSid', 'txHR', as.character(pol$id))]
    return(merged)
}
```

## Reproduce default analysis in Shiny app
The following code reproduces the default analysis in the Shiny app based on the example in the [introduction]() vignette.
```{r default}
exp.rate <- function(survival, year=5) -log(survival)/year

default <- list(pol=data.frame(num=c(1, 2, 3),
                               id=c('e0', 'e1', 'e4'),
                               name=c('Base Case',
                                      'Tamoxifen for ER+',
                                      'Tamoxifen for ER+ and 23% stage shift'),
                               pairnum=c(NA, NA, 2),
                               earlydetHR=c(1, 1, 0.77),
                               stringsAsFactors=FALSE),
                nh=compile_naturalhist(prop_adv=0.78,
                                       mortrates=c(Early=exp.rate(0.69),
                                                   Advanced=exp.rate(0.35)),
                                       subgroup_probs=c('ER+'=0.41, 'ER-'=1-0.41)))
default$map <- create_stageshift_map(default$nh)
default$tx <- with(default, create_treatment_distribution(nh, pol))
default$tx <- within(default$tx, {
                         txHR[grepl('ER[-]', SSid) & txSSid == 'Tamoxifen'] <- 1
                         e0[txSSid == 'None'] <- 0.8
                         e0[txSSid == 'Tamoxifen'] <- 0.2
                         e1[grepl('ER[+]', SSid) & txSSid == 'None'] <- 0
                         e1[grepl('ER[+]', SSid) & txSSid == 'Tamoxifen'] <- 1
                         e4[grepl('ER[+]', SSid) & txSSid == 'None'] <- 0
                         e4[grepl('ER[+]', SSid) & txSSid == 'Tamoxifen'] <- 1
                            })
plyr::d_ply(default$tx,
            'SSid',
            with,
            stopifnot(sum(e0) == 1 &
                      sum(e1) == 1 &
                      sum(e4) == 1))
set.seed(98103)
uganda_default <- with(default, simpolicies(pol, nh, tx,
                                            minage=30,
                                            maxage=49,
                                            futimes=c(10, 20),
                                            popsize=100000,
                                            sims=100))
knitr::kable(uganda_default[['10']])
knitr::kable(uganda_default[['20']])
```

Questions:

* Should treatment distributions sum to 1 within each stage-subgroup (as in
  [introduction]()) or sum to 1 within each stage (as in Shiny app)?

* Why does tx input give tamoxifen txHR=0.7 even for ER- cancers in Shiny app?

## Incorporate HER2 subgroups and add new policy scenario
The following code extends subgroups assuming 15% of cancers are HER2+, uncorrelated with ER status and stage at presentation, based on the example in the [introduction]() vignette. It also assumes efficacy of trastuzumab implies HR=0.60
```{r her2}
her2 <- list(pol=data.frame(num=c(1, 2, 3, 4),
                            id=c('e0', 'e1', 'e4', 'e10'),
                            name=c('Base Case',
                                   'Tamoxifen for ER+',
                                   'Tamoxifen for ER+ and 23% stage shift',
                                   'Tamoxifen for ER+, trastuzumab for HER2+, and 23% stage shift'),
                            pairnum=c(NA, NA, 2, 3),
                            earlydetHR=c(1, 1, 0.77, 0.77),
                            stringsAsFactors=FALSE),
             nh=compile_naturalhist(prop_adv=0.78,
                                    mortrates=c(Early=exp.rate(0.69),
                                                Advanced=exp.rate(0.35)),
                                    subgroup_probs=c('ER+HER2+'=0.41*0.15,
                                                     'ER-HER2+'=(1-0.41)*0.15,
                                                     'ER+HER2-'=0.41*(1-0.15),
                                                     'ER-HER2-'=(1-0.41)*(1-0.15))))
her2$map <- create_stageshift_map(her2$nh)
her2$tx <- with(her2, create_treatment_distribution(nh, pol,
                                                    treat_hrs=c('None'=1,
                                                                'Tamoxifen'=0.7,
                                                                'Trastuzumab'=0.6),
                                                    treat_probs=c('None'=1,
                                                                  'Tamoxifen'=0,
                                                                  'Trastuzumab'=0)))
her2$tx <- within(her2$tx, {
                         txHR[grepl('ER[-]', SSid) & txSSid == 'Tamoxifen'] <- 1
                         txHR[grepl('HER2[-]', SSid) & txSSid == 'Trastuzumab'] <- 1
                         e1[grepl('ER[+]', SSid) & txSSid == 'None'] <- 0
                         e1[grepl('ER[+]', SSid) & txSSid == 'Tamoxifen'] <- 1
                         e4[grepl('ER[+]', SSid) & txSSid == 'None'] <- 0
                         e4[grepl('ER[+]', SSid) & txSSid == 'Tamoxifen'] <- 1
                         e10[grepl('ER[+]', SSid) & txSSid == 'None'] <- 0
                         e10[grepl('ER[+]', SSid) & txSSid == 'Tamoxifen'] <- 1
                         e10[grepl('HER2[+]', SSid) & txSSid == 'None'] <- 0
                         e10[grepl('ER[+]HER2[+]', SSid) & txSSid == 'Tamoxifen'] <- 0
                         e10[grepl('HER2[+]', SSid) & txSSid == 'Trastuzumab'] <- 1
                            })
plyr::d_ply(her2$tx,
            'SSid',
            with,
            stopifnot(sum(e0) == 1 &
                      sum(e1) == 1 &
                      sum(e4) == 1 &
                      sum(e10) == 1))
set.seed(98103)
uganda_her2 <- with(her2, simpolicies(pol, nh, tx,
                                      minage=30,
                                      maxage=49,
                                      futimes=c(10, 20),
                                      popsize=10000))
knitr::kable(uganda_her2[['10']])
```

To do:

* Modify policies to use mammography (30% advanced stage at diagnosis) and chemotherapy for all ER- cancers and advanced ER+ cancers.

* Should trastuzumab be replaced with tamoxifen+trastuzumab so treatment distributions sum to 1 when stage-subgroups involve both treatments?

* Use wider age range? Use longer follow-up?

